---
title: 'Bariatric Whole Blood Data'
output: pdf_document
---

"Magrittr pipes and R assignment, use =,-> or <-!?"

R allows either ‘=’ or ‘->’ for assignment. Neither makes sense when using magrittr pipes and the dplyr functions. John Mount has proposed (http://www.win-vector.com/blog/2016/12/the-case-for-using-in-r/) that using the right arrow ‘->’ operator makes much more sense. As pipe operations become more common, this makes a lot of sense. 

Let's get the data first. This data came from a CIS project a student asked for help on. The help was not in the primary project, but just getting the data into R in a useable format. They were given the gene expression data in an SPSS save file.

```{r}
library(tidyverse)
library(haven)
Bariatric <- read_sav("BariatricWholeBloodDataSet.sav")
names(Bariatric)
```

```{r}
# Convert names to S.nnn.pre/post
gsub("Subject","S.",
     gsub("presurgery",".pre",
          gsub("postsurgery",".post",names(Bariatric))
     )
)
```
Use pipes
```{r}
names(Bariatric) %>%
  gsub("Subject","S.", . ) %>%
  gsub("presurgery",".pre", . ) %>%
  gsub("postsurgery",".post", . )
```

Now changing the names of the Bariatric data frame...
```{r, eval=FALSE, include=FALSE}
names(Bariatric) <-
  names(Bariatric) %>%
  gsub("Subject","S.", . ) %>%
  gsub("presurgery",".pre", . ) %>%
  gsub("postsurgery",".post", . )
```

That seems a bit tortured syntactically. But with the right assignment operator '->' we have the following, making it clear we're taking the names from Bariatric and replacing them with new ones.
```{r}
names(Bariatric) %>%
  gsub("Subject","S.", . ) %>%
  gsub("presurgery",".pre", . ) %>%
  gsub("postsurgery",".post", . ) ->
  names(Bariatric)

names(Bariatric)
```

Now let's try some data manipulaitons. We'll be using the following:
* gather:
* separate
* 

No pipes conversion
```{r}
# Gather into long form
  select(  
    separate(
        gather(Bariatric,
               key=id, value=response.var,
               S.004.pre:S.014.post),
        id,c("junk","subject","prepost")),
    -junk)
```

Using pipes is cleaner
```{r}
# Gather into long form and save
Bariatric.new <-
  Bariatric %>%
  gather(key=id, value=response.var,
         S.004.pre:S.014.post) %>%
  separate(id,c("junk","subject","prepost")) %>%
  select(-junk)
```

But that assignment doesn't flow with the pipes. Instead...
```{r}
# Gather into long form and save
Bariatric %>%
  gather(key=id, value=response.var,
         S.004.pre:S.014.post) %>%
  slice(1:10) %>%
  separate(id,c("junk","subject","prepost")) %>%
  select(-junk) -> 
  Bariatric.new
```
